Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Final State with Multivariate Techniques

### 2015-12-10 paper/sec-mcsamples.tex
polishing paper

Our modelling of detector simulation has been tuned
to lead to a mass resolution of
reconstructed Higgs candidates of around 10 GeV, consistent
with the hadronic mass resolutions of ATLAS and CMS~\cite{Aad:2012gxa,Chatrchyan:2013zna,Aad:2014xzb}.
%
In Fig.~\ref{fig:higgs-mass-resolution} we show the
invariant mass distributions of 
Higgs candidate in signal events for the
resolved category (see Sects.~\ref{sec:analysis}
and~\ref{sec:results}), together with
a Gaussian fit,
used to determine the mass resolution.
%
We indeed  find a mass resolution of about 8\%, corresponding
    to around 10 GeV for $m_h=125$ GeV.
%
    For completeness, in Fig.~\ref{fig:higgs-mass-resolution}
    we also show the corresponding
    mass distributions of Higgs candidates
    in the case of $\la n_{\rm PU}\ra=80$ PU events
    subtracted with {\tt SoftKiller} (see Sect.~\ref{sec:pileup}).
    %
    In the case of PU, the mass resolution is degraded down to
    around 11\%, corresponding to 
    14 GeV.
    %

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
    \vspace{-1cm}
  \includegraphics[width=0.49\textwidth]{plots/higgs_mass_res_noPU.pdf}
  \includegraphics[width=0.49\textwidth]{plots/higgs_mass_res_PU80.pdf}
  \caption{\small The invariant mass distributions of the leading
    Higgs candidates in signal events for the resolved category, without
    PU (left plot) and with PU (right plot),
    together with Gaussian fits  used to determine the mass resolution.
    %
    In the latter case, $\la n_{\rm PU}\ra=80$, with {\tt SoftKiller}
     used for PU subtraction.
}
\label{fig:higgs-mass-resolution}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

>>>>

Our modelling of detector simulation has been tuned
to lead to a mass resolution of
the reconstructed Higgs candidates of around 10 GeV, consistent
with the hadronic mass resolutions of the
ATLAS and CMS detectors~\cite{Aad:2012gxa,Chatrchyan:2013zna,Aad:2014xzb}.
%
By fitting a Gaussian to the 
invariant Higgs mass distributions of 
in signal events, we can 
determine the effective mass resolution
in the present analysis.
%
We find a resolution of about 8\% in the resolved
category, corresponding to a window
of around 10 GeV for $m_h=125$ GeV.
%
Fitting the Higgs    mass distributions 
in the case of $\la n_{\rm PU}\ra=80$ PU events
with subtraction
(see Sect.~\ref{sec:pileup}),
the mass resolution is degraded down to
    around 11\%, corresponding to a windown of
    14 GeV around the Higgs mass.
    %

### 2015-12-10 paper/sec-mva.tex
polishing paper

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.48\textwidth]{plots/ssb_noPU.pdf}
\includegraphics[width=0.48\textwidth]{plots/sb_noPU.pdf}
\caption{\small
  The values of the signal significance, $S/\sqrt{B}$, and of the
  signal over background ratio, $S/B$, for the boosted, intermediate
  and resolved categories as  function of the cut
  $y_{\rm cut}$ in the ANN output.
  %
  The $y_{\rm cut}=0$
  values correspond to the  cut-based analysis
  results from Table~\ref{tab:cutflow_noPU_1}.
}
\label{fig:sb_mva}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

The results for the signal significance $S/\sqrt{B}$ and
the signal over background ratio
$S/B$ as a function of $y_{\rm cut}$
for the three categories are
Fig.~\ref{fig:sb_mva}.
%
The values 
for $y_{\rm cut}=0$ correspond to those
the end of the cut-based analysis, see
Table~\ref{tab:cutflow_noPU_1}.
%
We observe how in the three
 categories there is a marked  improvement in signal
significance as compared to the pre-MVA results.
%
We also observe a substantial enhancement in $S/B$, arising
from the background suppression indeed by the MVA, reaching
values of 1\%, 6\% and 3.5\% in the resolved,
intermediate and boosted categories.
%
This improvement in $S/B$ is crucial to ensure the feasibility
of this measurement, since it allows systematic
uncertainties in the background determination to
be at the same level of a few percent level.

@@

In this section we study how the MVA results are modified when
signal and background events are embedded in a high-PU environment.
%
The simulation of PU and the corresponding subtraction with
{\tt SoftKiller} have been performed as described
in Sect.~\ref{sec:pileup}.
%
Following this, the cut-based analysis and the subsequent
MVA optimization have been applied using the same
settings as in the case without PU.
%
In Table~\ref{tab:cutflow_SKPU80_1} we provide the results
of the cut-based analysis cut-flow, for the case with $\la n_{\rm PU}\ra=80$+SK.
%
It is clear that the pre-MVA 
signal significance have been seriously degraded
as compared to the case without PU.
%
We now find values for $S/\sqrt{B}$ of 0.4, 0.3 and 0.9, in the resolved,
intermediate and boosted categories, respectively, to be compared
with the corresponding values without PU, namely 1.6, 1.6 and 2.7.
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \scriptsize
  \input{tables/table_SKPU80_1.tex}
  $\,$ \\
  \vspace{0.5cm}
  \input{tables/table_SKPU80_2.tex}
  $\,$ \\
  \vspace{0.5cm}
    \input{tables/table_SKPU80_3.tex}
    \caption{\small
      Same as Table~\ref{tab:cutflow_noPU_1}, for the analysis
      including PU with $\la n_{\rm PU}\ra=80$ and SK subtraction.
      \label{tab:cutflow_SKPU80_1}}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Following the cut-based analysis, the surviving signal and background
MC events in the case of PU are used to re-train the MVA.
%
In Table~\ref{table:cutflowMVA_PU}
we provide  the number of signal and
    background events expected
    at the HL-LHC with $\mathcal{L}=3$ ab$^{-1}$,
    as well as the
    corresponding values for $S/\sqrt{B}$ and $S/B$,
    for both the cut-based analysis (corresponding
    to $y_{\rm cut}=0$) and after the
    optimal MVA cut, for the simulations
    with $\la n_{\rm PU}\ra=80$
    and SK subtraction.
    %
    From Table~\ref{table:cutflowMVA_PU} we observe that, thanks
to the MVA, the signal significance and the
signal over background ratio are improved from 0.35 and $4\cdot 10^{-4}$
(0.25 and $1\cdot 10^{-4}$) up to 2.9 and 0.034 (2.5 and 0.15)
in the boosted (resolved) category.
%
The intermediate category, on the other hand, exhibits a rather lower post-MVA value of $S/\sqrt{B}$, around
1.1.
%
We thus conclude that, as was already the case
without PU,
the boosted category is the most promising
one for the study of Higgs pair production in the $b\bar{b}b\bar{b}$
final state
at the HL-LHC, also for the case of a high-PU environment, and that
important information will also be obtained from
the resolved category, which has a comparable significance.

@@

Another important result from Table~\ref{table:cutflowMVA_PU} is that,
for the optimal MVA cut, the signal over background ratio
is not unfeasibly small: we obtain that $S/B$ is 3.4\% (1.5\%)
in the boosted (resolved) categories.
%
This is indicates that while this measurement is still very challenging,
requiring a careful extraction from the data of the QCD
background, it is certainty within reach.
%
As in the case without PU, 
this analysis  indicates that a substantial increase in both
$S/B$ and $S/\sqrt{B}$ could be achieved by reducing the background component
that arises from the mis-identification of light and charm
fakes, compare for in particular the columns $S/\sqrt{B_{\rm tot}}$
and  $S/\sqrt{B_{\rm 4b}}$ in  Table~\ref{tab:cutflow_SKPU80_1}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.48\textwidth]{plots/ssb_noPU.pdf}
\includegraphics[width=0.48\textwidth]{plots/sb_noPU.pdf}
\caption{\small
  The values of the signal significance, $S/\sqrt{B}$, and of the
  signal over background ratio, $S/B$, for the boosted, intermediate
  and resolved categories as  function of the cut
  $y_{\rm cut}$ in the ANN output.
  %
  The $y_{\rm cut}=0$
  results are those at the end of the cut-based
  analysis.
}
\label{fig:sb_mva}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

The results for the signal significance $S/\sqrt{B}$ and
the signal over background ratio
$S/B$ as a function of $y_{\rm cut}$
for the three categories are
Fig.~\ref{fig:sb_mva}.
%
The values 
for $y_{\rm cut}=0$ correspond to those at
the end of the cut-based analysis.
%
We observe how in the three
 categories there is a marked  improvement in signal
significance as compared to the pre-MVA results.
%
We also observe a substantial enhancement in $S/B$, arising
from the background suppression indeed by the MVA, reaching
values of 1\%, 6\% and 3.5\% in the resolved,
intermediate and boosted categories.
%
This improvement in $S/B$ is crucial to ensure the feasibility
of this measurement, since it allows systematic
uncertainties in the background determination to
be at the same level of a few percent level.

@@

In this section we study how the MVA results are modified when
signal and background events are embedded in a high-PU environment.
%
The simulation of PU and the corresponding subtraction with
{\tt SoftKiller} have been performed as described
in Sect.~\ref{sec:pileup}.
%
Following this, the cut-based analysis and the subsequent
MVA optimization have been applied using the same
settings as in the case without PU.
%
In Table~\ref{table:cutflowMVA_PU} we provide the results
  at the end of the cut-based analysis,
for the case with $\la n_{\rm PU}\ra=80$+SK.
%
It is clear that the pre-MVA 
signal significance have been seriously degraded
as compared to the case without PU.
%
We now find values for $S/\sqrt{B}$ of 0.4, 0.3 and 0.9, in the resolved,
intermediate and boosted categories, respectively, to be compared
with the corresponding values without PU, namely 1.6, 1.6 and 2.7.
%

Subsequently to 
 the cut-based analysis, the surviving signal and background
MC events in the case of PU are used to re-train the MVA.
%
In Table~\ref{table:cutflowMVA_PU}
we also provide  the number of signal and
    background events expected
    at the HL-LHC with $\mathcal{L}=3$ ab$^{-1}$,
    as well as the
    corresponding values for $S/\sqrt{B}$ and $S/B$,
    for both the cut-based analysis (corresponding
    to $y_{\rm cut}=0$) and after the
    optimal MVA cut, for the simulations
    with $\la n_{\rm PU}\ra=80$
    and SK subtraction.
    %
    From Table~\ref{table:cutflowMVA_PU} we observe that, thanks
to the MVA, the signal significance and the
signal over background ratio are improved from 0.35 and $4\cdot 10^{-4}$
(0.25 and $1\cdot 10^{-4}$) up to 2.9 and 0.034 (2.5 and 0.15)
in the boosted (resolved) category.
%
The intermediate category, on the other hand, exhibits a rather lower post-MVA value of $S/\sqrt{B}$, around
1.1.
%
We thus conclude that, as was already the case
without PU,
the boosted category is the most promising
one for the study of Higgs pair production in the $b\bar{b}b\bar{b}$
final state
at the HL-LHC, also for the case of a high-PU environment, and that
important information will also be obtained from
the resolved category, which has a comparable significance.

@@

Another important result from Table~\ref{table:cutflowMVA_PU} is that,
for the optimal MVA cut, the signal over background ratio
is not unfeasibly small: we obtain that $S/B$ is 3.4\% (1.5\%)
in the boosted (resolved) categories.
%
This is indicates that while this measurement is still very challenging,
requiring a careful extraction from the data of the QCD
background, it is certainty within reach.
%
As in the case without PU, 
this analysis  indicates that a substantial increase in both
$S/B$ and $S/\sqrt{B}$ could be achieved by reducing the background component
that arises from the mis-identification of light and charm
fakes.
